Implications of animal movement rate variation in spatial patterns of seed dispersal

### 2021-08-16 Ch1_movement_rates/Ch1_markdown.Rmd
populations with different levels of variation in the movement rates

@@

```{r}
mycols <- c("black", "blue", "purple")
#create density plots
curve(dlnorm(x, meanlog=1, sdlog=.3), from=0, to=10, col=mycols[1],
      main = 'Log Normal Distribution', #add title
      ylab = 'Density', #change y-axis label
      )
curve(dlnorm(x, meanlog=1, sdlog=.5), from=0, to=10, col=mycols[2], add=TRUE)
curve(dlnorm(x, meanlog=1, sdlog=1), from=0, to=10, col=mycols[3], add=TRUE)

@@

Now, lets sample a movement rate, $m$, for 10 individuals $(i.e. j = 1...10)$ from the first population $(i.e. m_{1,j} = m_{1,1}, ....., m_{1,10})$. We will simulate animal movement with a simple random walk, with movement length $l$ sampled from an exponential distribution, $l \sim exp(m_{i,j})$, movement angle, $\alpha$, from a uniform distribution, $\alpha \sim U(0, 360)$. We can visualize the movement for one of the individuals in the first population:

```{r}
m_1 <- rlnorm(10, meanlog = 1, sdlog = 0.3)
m_2 <- rlnorm(10, meanlog = 1, sdlog = 0.5)
m_3 <- rlnorm(10, meanlog = 1, sdlog = 1)
```

@@

@@

```{r}
t <- 1000 # Number of time steps

tru.rate <- m_1[1]
movedist <- rexp(t, rate = tru.rate)
angle <- runif(t, min = 0, max = 360)
distx <- movedist*cos(angle)
xloc <- c(0, cumsum(distx))
disty <- movedist*sin(angle)
yloc <- c(0, cumsum(disty))
animalTraj <- data.frame(time = 0:t, xloc = xloc, yloc = yloc)

plot(x = xloc, y = yloc, type = "l", main = paste("Rate", tru.rate))

```{r}
par(mfrow=c(2,5))

for(j in 1:10){
  sim_movement(m_1[j])}

Use ggplot to make multiple rounds of each individual
```{r}
library(dplyr)
library(ggplot2)

@@

```
Overlay 5 runs for that individual

@@

for(k in 1:8){
  a <- sim_movement(m_1[1], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, color = run)) +
  geom_path()
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

@@

```{r}
mycols <- c("black", "blue", "purple")
#create density plots
curve(dlnorm(x, meanlog=1, sdlog=.3), from=0, to=10, col=mycols[1],
      main = 'Log Normally distributed movement rates for three populations', #add title
      ylab = 'Density', #change y-axis label
      )
curve(dlnorm(x, meanlog=1, sdlog=.5), from=0, to=10, col=mycols[2], add=TRUE)
curve(dlnorm(x, meanlog=1, sdlog=1), from=0, to=10, col=mycols[3], add=TRUE)

@@

Now, lets sample a movement rate, $m$, for 5 individuals $(i.e. j = 1, ...,5)$ from the first population $(i.e. m_{1,j} = m_{1,1}, ....., m_{1,5})$. We will simulate animal movement with a simple random walk, with movement length $l$ sampled from an exponential distribution, $l \sim exp(m_{i,j})$, movement angle, $\theta$, from a uniform distribution, $\theta \sim U(0, 360)$. We can visualize the movement for one of the individuals in the first population:

```{r}
m_1 <- sort(round(rlnorm(5, meanlog = 1, sdlog = 0.3),3))
m_2 <- sort(round(rlnorm(5, meanlog = 1, sdlog = 0.5),3))
m_3 <- sort(round(rlnorm(5, meanlog = 1, sdlog = 1),3))
```

@@

mycols <- rainbow(5, s=0.8)

# range <- c(-0.9,5)
#create density plots
curve(dexp(x, m_1[5]), 
      #from=range[1], to=range[2], 
      col=mycols[1],
      main = 'Exponentially distributed movement lengths for 5 individuals', #add title
      ylab = 'Density', #change y-axis label
      )
for(i in 1:4){
  curve(dexp(x, m_1[i]), 
        #from=range[1], to=range[2],
        col=mycols[i], add=TRUE)
}

#add legend
legend("topright", legend= m_1,
       col=mycols, lty=1, cex=1.2)
```

```{r}
t <- 100 # Number of time steps
individual <- 5

tru.rate <- m_1[individual] # In case I want to do scale 1/rate
movedist <- rexp(t, rate = tru.rate)
angle <- runif(t, min = 0, max = 360)
distx <- movedist*cos(angle)
xloc <- c(0, cumsum(distx))
disty <- movedist*sin(angle)
yloc <- c(0, cumsum(disty))
animalTraj <- data.frame(time = 0:t, xloc = xloc, yloc = yloc)

plot(x = xloc, y = yloc, col = mycols[individual], type = "l", main = paste("Rate", tru.rate))

```{r eval=FALSE}
par(mfrow=c(2,5))

for(j in 1:individual){
  sim_movement(m_1[j])}

@@

Use ggplot to make multiple rounds of each individual
```{r eval=FALSE}

@@

Population1
```{r}

@@

for(j in 1:individual){
  for(k in 1:10){
  a <- sim_movement(m_1[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw()
```

Population2
```{r}

for(j in 1:individual){
  for(k in 1:10){
  a <- sim_movement(m_2[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw()
```

Population3
```{r}

for(j in 1:individual){
  for(k in 1:10){
  a <- sim_movement(m_3[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw()
```
